---
title: "Gapminder Analysis"
author: "Julie Cardon"
date: "2024-02-07"
output: 
  html_document:
    toc: yes
    toc_float: yes
    code_folding: hide
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

(The above chunk determines whether or not chunks, messages, etc. will
be shown-- can set width and height of figures, where they're being
saved, etc.) In this document I will analyze how life expectancy has
changed over time (include purpose of document) html is nice because it
has continuous scroll (no page breaks, good for big docs) to put r code
into document, put in chunks (title goes in curly brackets, can put in
header as well) document and console can use different working
directories, use knit directory to fix we used the settings button to
remove output from document (see in console- you can choose) \# Reading
in gapminder data (header has hashtag and space before)

```{r reading-in-data}
gapminder_df <- read.csv("data/gapminder_data.csv")

```

## A smaller header

### An even smaller header

###### A very small header

Lists: make with asterisk and a space

-   list item 1
-   list item 2

Make hyphen lists: - Hyphen list - Hyphen list

Numbered lists: 1. Number one 2. Number two \* Add 4 spaces for a sub
list 3. Number 3 Put a space between header and list

We make text **bold with two asterisks** or *italics with one asterisk*

We can also insert greek letters and equations with $\mu$ or $\Delta$

We can make hyperlinks with square brackets for word we want linked,
followed by parenthesis around the link like
[hyperlinks](https://pad.carpentries.org/Spring2024-BIOMI6300)

# Calculating summary statistics

how many observations, and the median echo=FALSE within the chunk title
won't show the code in the document

```{r calculating-median-pop, echo=FALSE}
med_pop <- median(gapminder_df$pop)
num_obs <- nrow(gapminder_df)
```

As shown above, the median population was 7,023,596

Over `r num_obs` observations, the median population was
`r round(med_pop/10e5, 3)` million people. notice in-line r codes, and
how that shows up in the knitted text.

#Plotting life Expectancy Over the Years

```{r line-graph, warning = FALSE}
library(ggplot2) #(the hashtag makes it so r doesn't read the code)
ggplot(gapminder_df, aes(x = year, y = lifeExp, group = country, color = continent)) + geom_line() + scale_y_log10() + scale_x_continuous(limits = c(1997, 2007))
```

warnings can be silenced in title of code chunk \# Make a line plot \#
log scale y axis \# Bonus: Change the theme

always have the library loading step in your r chunk eval = false (shows
code, but doesn't run it, output = false, run code but not show)

#Plotting Population Distribution

```{r Histogram, message = FALSE, fig.width=10}
ggplot(gapminder_df, aes(x=pop)) + geom_histogram()
```

message in r- nothing went wrong, but you should know... (message is not
useful to reader, so you can put in message = false so it doesn't print
the message)

YAML is the program that makes the header (can add options to html)

toc gives us a table of contents

code_folding allows user to choose to show the code

You can use the visual view to work like in a word processor and to
insert tables

+---------------------+----------------------+----------------------+
| Col1                | Col2                 | Col3                 |
+=====================+======================+======================+
| put some stuff here | but it looks weird   | you can insert       |
|                     | in markdown          | tables as an image   |
|                     |                      | instead and they'll  |
|                     |                      | be in your image     |
|                     |                      | folder               |
+---------------------+----------------------+----------------------+
|                     |                      |                      |
+---------------------+----------------------+----------------------+
|                     |                      |                      |
+---------------------+----------------------+----------------------+
any code can be used within an R markdown document, like python and bash
